Report on the Third Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE3)

### 2016-08-31 WSSSPE3/report/WSSSPE3_review.tex
small fix in 5 intro, removing first person in 5.1 for group stuf

After the keynote and lightning talks, the workshop facilitators led an exercise in which the
attendees first suggested additional topics beyond those in the call for participation (as listed in
\S\ref{sec:preworkshop}).  One topic that was introduced at this point was ``Building Sustainable
User Communities for Scientific Software.''  The full set of topics were then placed on
flipchart-sized pieces of paper around the walls, and the attendees voted on which topics
they were strongly interested in working on at WSSSPE3, and which topics they were 
generally interested in contributing to, but not as strongly.  This led to some topics being
taken out of the mix for the rest of the workshop, since not enough people wanted to contribute
to them to lead to a useful discussion.  Additionally, some topics were brought together by the
participants who felt they were closely linked and that a group could address multiple of them
in a single discussion.

>>>>

After the keynote and lightning talks, the workshop facilitators led an exercise to create working
groups for the rest of the meeting.  The
attendees first suggested additional topics beyond those in the call for participation (as listed in
\S\ref{sec:preworkshop}).  One topic that was introduced at this point was ``Building Sustainable
User Communities for Scientific Software.''  The full set of topics were then placed on
flipchart-sized pieces of paper around the walls, and the attendees voted on which topics
they were strongly interested in working on at WSSSPE3, and which topics they were 
generally interested in contributing to, but not as strongly.  This led to some topics being
taken out of the mix for the rest of the workshop, since not enough people wanted to contribute
to them to lead to a useful discussion.  Additionally, some topics were brought together by the
participants who felt they were closely linked and that a group could address multiple of them
in a single discussion.

### 2016-08-31 WSSSPE3/report/wg_main_best_practices.tex
small fix in 5 intro, removing first person in 5.1 for group stuf

The creation of such a ``best practices'' document will build upon the range of
activities and topics discussed at WSSSPE3 and associated prior meetings. We
will attempt to distill the emerging body of knowledge into this document. The
large number of articles from the NSF funded SI2 projects (SSE and SSI),
``lightning talks'', ``white papers,'' and reports from different workshops have
created a large if somewhat diffuse source for this report.

@@

Core questions that will need to be explored are in reproducibility, 
reliability,  usability, extensibility, knowledge management, as well as continuity
(transitions between people). Answers to these questions will guide us on how a software
tool becomes part of the core workflow of well identified users (stakeholders)
relating to tool success and hence sustainability.
%\katznote{prev sentence is complex and awkward} % see fix
Ideas  that may need to be explored include:
\begin{itemize}
\item Requirements engineering to create tools with immediate uptake;
\item When should software ``die''?
\item Catering to disruptive developments in environment (e.g., new hardware,
new methodology);
\item Dimensions of sustainability: economic, technical, environmental and
obsolescence. 
%\katznote{not sure what the prev. comment goes with} %social.
\end{itemize}

@@

\subsubsection{Plans}
%\todo{short text here - not bullets}
The creation of a document on best practices needs a large and diverse community
involved. We have enlisted over ten contributors from the attendees at the
WSSSPE3 and those on the mailing list. The primary mechanism for developing this
document will be to examine and analyze the success of several well known
community scientific software and organizations supporting scientific software.
We will attempt then to abstract general principles and best practices. Some of
the tools identified for such analysis are the general purpose PETSc toolkit for
linear system solution, NWChem for computational chemistry and the CIG
(Computational Infrastructure for Geodynamics) organization dedicated to
supporting an ensemble of related tools for the geodynamics community. We also
established a timeline and a rough outline (see
Appendix~\ref{sec:appendix_best_practices}) for the report.

>>>>

The creation of such a ``best practices'' document will build upon the range of
activities and topics discussed at WSSSPE3 and associated prior meetings. This
working group
will attempt to distill the emerging body of knowledge into this document. The
large number of articles from the NSF funded SI2 projects (SSE and SSI),
``lightning talks,'' ``white papers,'' and reports from different workshops have
created a large if somewhat diffuse source for this report.

@@

Core questions that will need to be explored are in reproducibility, 
reliability,  usability, extensibility, knowledge management, as well as continuity
(transitions between people). Answers to these questions will guide the group
to learn how a software
tool becomes part of the core workflow of well-identified users (stakeholders)
relating to tool success and hence sustainability.
%\katznote{prev sentence is complex and awkward} % see fix
Ideas  that may need to be explored include:
\begin{itemize}
\item Requirements engineering to create tools with immediate uptake;
\item When should software ``die''?
\item Catering to disruptive developments in environment (e.g., new hardware,
new methodology);
\item Dimensions of sustainability: economic, technical, environmental and
obsolescence. 
%\katznote{not sure what the prev. comment goes with} %social.
\end{itemize}

@@

\subsubsection{Plans}
%\todo{short text here - not bullets}
The creation of a document on best practices needs a large and diverse community
involved. The group has enlisted over ten contributors from the attendees at the
WSSSPE3 and those on the mailing list. The primary mechanism for developing this
document will be to examine and analyze the success of several well known
community scientific software and organizations supporting scientific software.
The group will attempt then to abstract general principles and best practices. Some of
the tools identified for such analysis are the general purpose PETSc toolkit for
linear system solution, NWChem for computational chemistry and the CIG
(Computational Infrastructure for Geodynamics) organization dedicated to
supporting an ensemble of related tools for the geodynamics community. The group also
established a timeline and a rough outline (see
Appendix~\ref{sec:appendix_best_practices}) for the report.
